Let There Be Light: \\ Early Observations of the Normal Type Ia Supernova \abc\ \\Show No Evidence for a Dark Period

### 2017-07-28 doc/ms.tex
Update text in the asbtract of the manuscript to reflect new conclusions.

Early observations of type Ia supernovae (SNe) provide a unique probe of their progenitor systems and explosion physics. Here, we report the intermediate Palomar Transient Factory (iPTF) discovery of an extraordinarily young Type Ia supernova, \abc. The initial detection of the SN was made only $0.18$~d after the time of first light. In the $\sim$24 hr after discovery, \abc\ rose by $\sim$2 mag in the $g_\mathrm{PTF}$ band, which is very rare for young type Ia SNe. Furthermore, by measuring the evolution of the photospheric velocity we determine that the time of explosion for \abc\ is approximately equal to the time of first light. Spectra obtained in the week after explosion reveal the presence of strong \ion{C}{2} absorption, which steadily fades over that time period. We show that the unusually fast, near-linear rise in flux for \abc\ cannot be explained by either SN shock breakout and the associated subsequent cooling or the collision of the SN ejecta with a stellar companion. Instead, we show each of the unusual characteristics of \abc: (i) the rapid, near-linear rise, (ii) the lack of a dark period between explosion and first light, and (iii) the strong lines from ionized carbon in the earliest spectra, can all be explained if radioactive $^{56}$Ni is strongly mixed throughout the SN ejecta. In the next few years, dozens of very young \textit{normal} SNe Ia will be discovered, and observations similar to those presented here will constrain the white dwarf explosion mechanism.

>>>>

Early observations of Type Ia supernovae (SNe) provide a unique probe of their progenitor systems and explosion physics. Here, we report the intermediate Palomar Transient Factory (iPTF) discovery of an extraordinarily young SN Ia, \abc. The initial detection was made only $0.18\pm^{0.07}_{0.15}$~d after SN first light. In the $\sim$24 hr after discovery, \abc\ rose by $\sim$2 mag, following a near-linear rise in flux for $\sim$4 d. Assuming the photospheric velocity follows a $t^{-0.22}$ power law, the time of explosion for \abc\ is approximately equal to the time of first light. Strong \ion{C}{2} absorption is detected in the early spectra of \abc, before disappearing after $\sim$7~d. Unlike SN\,2011fe, the $(B-V)_0$ colors of \abc\ are blue and nearly constant in the days after explosion. We show that the fast, near-linear rise in flux for \abc\ cannot be explained by either SN shock breakout and the associated, subsequent cooling or the collision of the SN ejecta with a stellar companion. Instead, we argue that the early characteristics of \abc, including: (i) the rapid, near-linear rise, (ii) the non-evolving blue colors, and (iii) the strong lines from ionized carbon in the earliest spectra, are the result of either vigorous mixing of radioactive $^{56}$Ni in the SN ejecta, or the interaction of the ejecta with diffuse material, or a combination of the two. In the next few years, dozens of very young \textit{normal} SNe Ia will be discovered, and observations similar to those presented here will constrain the white dwarf explosion mechanism.
